Comparison of delivering limited fault tolerance via publish-subscribe messaging and message queuing in asynchronous distributed systems

### 2018-11-10 latex/article.tex
tables n stuff

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{blindtext}
\usepackage{textcomp}
\usepackage{float}
\usepackage{xcolor}
\usepackage{stfloats}
\usepackage{enumitem}
\usepackage{titlesec}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{distsys.png}
    \caption{General distributed system composition with one communication intermediary.}
\end{figure}

@@

\begin{itemize}
    \item Space decoupling: Publishers and subscribers do not have to know each other.
    \item Time decoupling: Subscriber can be offline at the time of publishing, publisher can be offline when subscriber receives the message.
    \item Synchronization decoupling: Publishers can broadcast a message and forget about it, subscribers are able to receive messages and handle them on their own pace.
\end{itemize}
With these definitions we can express the publish-subscribe messaging system with the diagram presented in Figure 2.
\begin{figure}
    \centering
    \includegraphics[scale=0.45]{PubSubTheo.png}
    \caption{Publish-subscribe messaging system.}
\end{figure}

@@

With these definitions we can express the message queue with the diagram presented in Figure 3.
\begin{figure}
    \centering
    \includegraphics[scale=0.45]{MQTheo.png}
    \caption{Message queue}
\end{figure}

@@

\section{Results}
\subsection{First perspective: qualitative analysis}
\subsubsection{Publish-subscribe messaging system}
The core advantage of publish-subscribe messaging system comes from the three dimensions of decoupling. They occur stronger than in message queues, which makes publish-subscribe better suited for loosely coupled systems. Main selling point of the loose coupling is that new services can be added to listen on pre-existing topics, and this does not require any changes in the rest of the system. We have also found this to be extremely beneficial in our advertisement network, where we can trial and launch new features without risking bugs in current production software.

@@

Last advantage of publish-subscribe is that if multiple services are subscribed to the same message, we must only store the message content once, whereas message queue requires either duplication of the message, or increased complexity and computational requirements to store only one copy. When the amount of subscribers wishing to read similar messages grows large, message queue would run out of memory faster, thus leading to increased operational costs or higher latencies.

@@

Message queue also makes subscriber services more robust since they do not need to expose themselves in the network. This prevents adversaries from attempting to flood the service with bogus requests, and somewhat reduces the complexity of the software. These advantages leave the developers more time to focus on the actual features of the service.

\subsection{Second perspective: performance analysis}
\subsubsection{Continuous streaming}
Continuous streaming in an environment where all the publishers and subscribers are online is perhaps the most common use case for an event system. This test tries to simulate such an environment, and yields results that allow us to make some conclusion on the comparison  between publish-subscribe and message queue.

In our first benchmark we can see that when we make no assumptions on the performance of the subscribers, publish-subscribe tends to work quite fast if subscribers can keep up with the workload, whereas message queue delivers slower message handling when there is no logic increasing the amount of messages fetched, even if it could handle more traffic.

The same observations hold for larger amount of data, which leads to the conclusion that while continuously streaming in a rather simple distributed system, publish-subscribe fares better.

\begin{figure}
    \centering
    \includegraphics[scale=0.30]{ContStream24k.png}
    \caption{24000 messages with 5 millisecond interval, Y-axis is in seconds.}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=0.30]{ContStream96k.png}
    \caption{96000 messages with 5 millisecond interval, Y-axis is in seconds.}
\end{figure}

@@

As the figure shows, message queue subscribers are able to catch up somewhat faster when there are multiple messages pending to be processed, since publish-subscribe messaging system is constrained due to not knowing how much traffic the subscribers could handle. There are possible implementations of publish subscribe that would allow adjusting of the publishing rate, but those fall outside the scope of comparing the approaches on this paper's purpose in delivering a general comparison.

With larger message set the difference becomes even more pronounced in terms of seconds but the ratio stays roughly the same. For failure recovery scenarios it seems that message queues would fare better especially when no assumptions on the workings of the rest of the system can be made.

\begin{figure}
    \centering
    \includegraphics[scale=0.30]{ColdBoot24k.png}
    \caption{24000 messages with 5 millisecond interval, Y-axis is in seconds.}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=0.30]{ColdBoot96k.png}
    \caption{96000 messages with 5 millisecond interval, Y-axis is in seconds.}
\end{figure}

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{blindtext}
\usepackage{textcomp}
\usepackage{float}
\usepackage{xcolor}
\usepackage{stfloats}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{booktabs}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.6]{distsys.png}
    \caption{General distributed system composition with one communication intermediary.}
\end{figure}

@@

\begin{itemize}
    \item Space decoupling: Publishers and subscribers do not have to know each other.
    \item Time decoupling: Subscriber can be offline at the time of publishing, publisher can be offline when subscriber receives the message.
    \item Synchronization decoupling: Publishers can broadcast a message and forget about it, subscribers are able to receive messages and handle them on their own pace.
\end{itemize}
With these definitions we can express the publish-subscribe messaging system with the diagram presented in Figure 2.
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.45]{PubSubTheo.png}
    \caption{Publish-subscribe messaging system.}
\end{figure}

@@

With these definitions we can express the message queue with the diagram presented in Figure 3.
\begin{figure}[h]
    \centering
    \includegraphics[scale=0.45]{MQTheo.png}
    \caption{Message queue}
\end{figure}

@@

\section{Results}
\subsection{First perspective: qualitative comparison}
In this section we argue for the benefits of each system that can be derived from the previously presented definitions. A table will also be presented containing the main comparison points.

Qualitative comparison was constructed because we consider that performance is not always the most desirable function of an event system, i.e. certain features might be required by the business, or lower implementation complexity could be desired due to cost reasons REF TO SOFTWARE QUALITY.
\\

\subsubsection{Publish-subscribe messaging system}
The core advantage of publish-subscribe messaging system comes from the three dimensions of decoupling. They occur stronger than in message queues, which makes publish-subscribe better suited for loosely coupled systems. Main selling point of the loose coupling is that new services can be added to listen on pre-existing topics, and this does not require any changes in the rest of the system. We have also found this to be extremely beneficial in our advertisement network, where we can trial and launch new features without risking bugs in current production software.

@@

Last advantage of publish-subscribe is that if multiple services are subscribed to the same message, we must only store the message content once, whereas message queue requires either duplication of the message, or increased complexity and computational requirements to store only one copy. When the amount of subscribers wishing to read similar messages grows large, message queue would run out of memory faster, thus leading to increased operational costs or higher latencies.
\\

@@

Message queue also makes subscriber services more robust since they do not need to expose themselves in the network. This prevents adversaries from attempting to flood the service with bogus requests, and somewhat reduces the complexity of the software. These advantages leave the developers more time to focus on the actual features of the service.
\\

\subsubsection{Feature table}
\begin{table}[h]
\begin{tabular}{@{}lll@{}}
\toprule
Feature                                                                                                  & Publish-subscribe     & Message queue         \\ \midrule
Recovery from service failure                                                                            & \multicolumn{1}{c}{•} & \multicolumn{1}{c}{•} \\ \midrule
\begin{tabular}[c]{@{}l@{}}Adding subscribers without\\ modifications to rest of the system\end{tabular} & \multicolumn{1}{c}{•} &                       \\ \midrule
\begin{tabular}[c]{@{}l@{}}Allow subscribers to\\ control inbound message flow\end{tabular}              &                       & \multicolumn{1}{c}{•} \\ \midrule
\begin{tabular}[c]{@{}l@{}}Allow event system to\\ control outgoing message flow\end{tabular}            & \multicolumn{1}{c}{•} &                       \\ \midrule
\begin{tabular}[c]{@{}l@{}}Store identical messages\\ only once\end{tabular}                             & \multicolumn{1}{c}{•} &                       \\ \midrule
\begin{tabular}[c]{@{}l@{}}Secure communication\\ between publisher and subscriber\end{tabular}          &                       & \multicolumn{1}{c}{•} \\ \midrule
Non-exposed subscribers                                                                                  &                       & \multicolumn{1}{c}{•} \\ \bottomrule
\end{tabular}
\end{table}

\subsection{Second perspective: performance comparison}
In building an advertisement network that delivers content to over a billion devices, performance has been identified as a critical part of the system. Closely tied to general performance are lower latencies that have been observed to increase profits and advance customer satisfaction SOME REFS YO. Therefor having insight into performance differences is essential knowledge when business decisions are being made regarding the choice of an even system.
The following subsections will present concrete benchmarks and give a brief analysis on each one to better explain the differences.

\subsubsection{Continuous streaming}
Continuous streaming in an environment where all the publishers and subscribers are online is perhaps the most common use case for an event system. This test tries to simulate such an environment, and yields results that allow us to make some conclusion on the comparison  between publish-subscribe and message queue.

In our first benchmark (Figure 4) we can see that when we make no assumptions on the performance of the subscribers, publish-subscribe tends to work quite fast if subscribers can keep up with the workload, whereas message queue delivers slower message handling when there is no logic increasing the amount of messages fetched, even if it could handle more traffic.

The same observations hold for larger amount of data (Figure 5), which leads to the conclusion that while continuously streaming in a rather simple distributed system, publish-subscribe fares better.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.30]{ContStream24k.png}
    \caption{24000 messages with 5 millisecond interval, X-axis is in seconds.}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.30]{ContStream96k.png}
    \caption{96000 messages with 5 millisecond interval, X-axis is in seconds.}
\end{figure}

@@

As the Figure 6 shows, message queue subscribers are able to catch up somewhat faster when there are multiple messages pending to be processed, since publish-subscribe messaging system is constrained due to not knowing how much traffic the subscribers could handle. There are possible implementations of publish subscribe that would allow adjusting of the publishing rate, but those fall outside the scope of comparing the approaches on this paper's purpose in delivering a general comparison.

With a larger message set (Figure 7) the difference becomes even more pronounced in terms of seconds but the ratio stays roughly the same. For failure recovery scenarios it seems that message queues would fare better especially when no assumptions on the workings of the rest of the system can be made.

\begin{figure}
    \centering
    \includegraphics[scale=0.30]{ColdBoot24k.png}
    \caption{24000 messages with 5 millisecond interval, X-axis is in seconds.}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[scale=0.30]{ColdBoot96k.png}
    \caption{96000 messages with 5 millisecond interval, X-axis is in seconds.}
\end{figure}
